---
title: "Homework 11"
author: "Daniel Forcade"
date: "5/3/2021"
output:   
  rmdformats::readthedown:
  self_contained: true
thumbnails: true
lightbox: true
gallery: false
highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Library
library(tidyverse)
library(TeachingDemos)
library(ggpubr)
library(hrbrthemes)
library(ggsci)
library(psych)
```  

I'm taking the `mandible` dataset I was working with before, condensing the number of variables, and creating eight batch files of simulated data to compare.  
I will use mean, sd, and number of observations from the original dataset to create the eight batches.  

## Setup  

```{r set}
## Data Setup

## Set seed
set.seed(77)

## Read in CSV
mandible <- read.csv('mandible.csv')

## Remove excess columns
mandible <- mandible[-c(3:ncol(mandible))]

# Summary
summary(mandible)

## Standard Dev
prothorax_sd <- sd(mandible$Prothorax.width)
mandible_sd <- sd(mandible$Mandible.length)

## Means
prothorax_mean <- mean(mandible$Prothorax.width)
mandible_mean <- mean(mandible$Mandible.length)

## Dimensions
obs <- nrow(mandible)
obs


```

## Filebuild  

```{r 1}
##################################################
# function: FileBuilder
# create a set of random files for regression
# input: fileN = number of files to create
#       : fileFolder = name of folder for random files
#       : fileSize = c(min,max) number of rows in file
#       : fileNA = number on average of NA values per column
# output: set of random files
#------------------------------------------------- 
FileBuilder <- function(fileN=10,
                        fileFolder="mandibleFile/",
                        fileSize=c(15,100),
                        fileNA=3){
for (i in seq_len(fileN)) {
fileLength <- sample(fileSize[1]:fileSize[2],size=1) # get number of rows
Prothorax.width <- rnorm(n= obs, mean = prothorax_mean, sd = prothorax_sd) ## Create random Y values based on prothorax vector
Mandible.length <- rnorm(n = obs, mean = mandible_mean, sd = mandible_sd) ## Create random X values based on mandible vector
dF <- data.frame(Prothorax.width, Mandible.length) # bind into a data frame

# create label for file name with padded zeroes
fileLabel <- paste(fileFolder,
                       "mandible_batch",
                       formatC(i,
                       width=3,
                       format="d",
                       flag="0"),
                       ".csv",sep="")

# set up data file and incorporate time stamp and minimal metadata
write.table(cat("# Simulated Mandible data file for batch processing","\n",
                    "# timestamp: ",as.character(Sys.time()),"\n",
                    "# DCF","\n",
                    "# ------------------------", "\n",
                    "\n",
                    file=fileLabel,
                    row.names="",
                    col.names="",
                    sep=""))

# now add the data frame
write.table(x=dF,
            file=fileLabel,
            sep=",",
            row.names=FALSE,
            append=TRUE)


}
}
```  


```{r 2}
##################################################
# function: regStats
# fits linear model, extracts statistics
# input: 2-column data frame (x and y)
# output: slope, p-value, and r2
#------------------------------------------------- 
regStats <- function(d=NULL) {
             if(is.null(d)) {
               mandible_x <- runif(10)
               prothorax_y <- runif(10)
               d <- data.frame(Prothorax.width, Mandible.length)
             }
  . <- lm(data=d,d[,2]~d[,1])
  . <- summary(.)
  statsList <- list(Slope=.$coefficients[2,1],
                    pVal=.$coefficients[2,4],
                    r2=.$r.squared)
  return(statsList)

}
```

## Output  

```{r 3, message = FALSE, warning = FALSE}
#--------------------------------------------
# Global variables
fileFolder <- "mandibleFile/"
nFiles <- 8
fileOut <- "StatsSummary.csv"
#--------------------------------------------

# Create 100 random data sets
dir.create(fileFolder)
FileBuilder(fileN=nFiles)
fileNames <- list.files(path=fileFolder)

# Create data frame to hold file summary statistics
ID <- seq_along(fileNames)
fileName <- fileNames
slope <- rep(NA,nFiles)
pVal <- rep(NA,nFiles)
r2 <- rep(NA,nFiles)

statsOut <- data.frame(ID,fileName,slope,pVal,r2)

# batch process by looping through individual files
for (i in seq_along(fileNames)) {
  data <- read.table(file=paste(fileFolder,fileNames[i],sep=""),
                     sep=",",
                     header=TRUE) # read in next data file
  
  dClean <- data[complete.cases(data),] # get clean cases
  
  . <- regStats(dClean) # pull regression stats from clean file
  statsOut[i,3:5] <- unlist(.) # unlist, copy into last 3 columns
  
}
# set up output file and incorporate time stamp and minimal metadata
  write.table(cat("# Summary stats for ",
                    "batch processing of regression models","\n",
                    "# timestamp: ",as.character(Sys.time()),"\n",
                    "# DCF","\n",
                    "# ------------------------", "\n",
                    "\n",
                    file=fileOut,
                    row.names="",
                    col.names="",
                    sep=""))
  
# now add the data frame
  write.table(x=statsOut,
              file=fileOut,
              row.names=FALSE,
              col.names=TRUE,
              sep=",",
              append=TRUE)
  
print(statsOut)
```  

Great, that made eight .csv files/datasets with the parameters specified, and placed them all in the newly created file specified (\\ForcadeBio381\\HW11\\mandibleFile).  All of the slopes are fairly flat.  Let's see how this compares visually against our original dataset.  



## Investigate Batch vs Data  

```{r explore}

#### Investigate Batch vs Data

## Load Data (Skip = 4 to clean out created metadata)
mandible_data <- read.csv('mandible.csv')
mandible_batch_1 <- read.csv('mandibleFile/mandible_batch001.csv', skip = 4)
mandible_batch_2 <- read.csv('mandibleFile/mandible_batch002.csv', skip = 4)
mandible_batch_3 <- read.csv('mandibleFile/mandible_batch003.csv', skip = 4)
mandible_batch_4 <- read.csv('mandibleFile/mandible_batch004.csv', skip = 4)
mandible_batch_5 <- read.csv('mandibleFile/mandible_batch005.csv', skip = 4)
mandible_batch_6 <- read.csv('mandibleFile/mandible_batch006.csv', skip = 4)
mandible_batch_7 <- read.csv('mandibleFile/mandible_batch007.csv', skip = 4)
mandible_batch_8 <- read.csv('mandibleFile/mandible_batch008.csv', skip = 4)

## Remove excess columns
mandible_data <- mandible_data[-c(3:ncol(mandible_data))]
```  


### Graph Set  

```{r graphs, warning = FALSE, message = FALSE}

### Graph Set
# Point Graph with lm

g0 <- ggplot(data = mandible_data,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm')+
  labs(title = "Mandible Data")

## Batch 1
g1 <- ggplot(data = mandible_batch_1,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm', color = 'red')+
  labs(title = "Batch 1")

## Batch 2
g2 <- ggplot(data = mandible_batch_2,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm', color = 'green')+
  labs(title = "Batch 2")

## Batch 3
g3 <- ggplot(data = mandible_batch_3,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm', color = 'yellow')+
  labs(title = "Batch 3")

## Batch 4
g4 <- ggplot(data = mandible_batch_4,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm', color = 'orange')+
  labs(title = "Batch 4")

## Batch 5
g5 <- ggplot(data = mandible_batch_5,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm', color = 'purple')+
  labs(title = "Batch 5")

## Batch 6
g6 <- ggplot(data = mandible_batch_6,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm', color = 'orchid4')+
  labs(title = "Batch 6")

## Batch 7
g7 <- ggplot(data = mandible_batch_7,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm', color = 'steelblue')+
  labs(title = "Batch 7")

## Batch 8
g8 <- ggplot(data = mandible_batch_8,
       mapping = aes(x = Prothorax.width, y = Mandible.length))+
  geom_point()+
  geom_smooth(method = 'lm', color = 'goldenrod4')+
  labs(title = "Batch 8")

## Print All
ggarrange(g0,g1,g2,g3,g4,g5,g6,g7,g8)

```  


### Pairs Panels  


```{r pairs}
## Pairs Panel Sampler Check

## OG Data
p0 <- pairs.panels(mandible_data)

## Batch 1
p1 <- pairs.panels(mandible_batch_1)

## Batch 2
p2 <- pairs.panels(mandible_batch_2)

## Batch 3
p3 <- pairs.panels(mandible_batch_3)

## Batch 4
p4 <- pairs.panels(mandible_batch_4)

## Batch 5
p5 <- pairs.panels(mandible_batch_5)
```  

All eight batch generations came out fairly similar; none are close to the correlation present in the original dataset.  The correlation exhibited between Mandible Size and Prothorax Width does not seem to be explained with tailored random generation.  





















