---
title: "Homework 8"
author: "Daniel Forcade"
date: "4/17/2021"
output:   
  rmdformats::readthedown:
  self_contained: true
thumbnails: true
lightbox: true
gallery: false
highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## Library
library(tidyverse)
library(psych)
library(dplyr)
library(reshape2)
library(ggsci)
library(ggthemes)
library(hrbrthemes)
library(ggpubr)
```

## Data Setup/Creation  
I used publicly available basketball metrics to create two distributions of `Assists per Game` separated by `Position`, for 2001 and 2021.  I used the average for mean, and a reasonable standard deviation based on basic research of what the expected variance would be.  
My hypothesis is that Assists per Game have increased in the last 10 years, with an emphasis on a more well-rounded game for each position.  This should result in higher Assists per Game averages for Power position players (`Centers`, `Power Forwards`) over the last 10 years.  


```{r datacreate}
set.seed(82)

## Two dataframes of related values
## Assists by position 2001 and 2021
## Mean/averages calculated from data on basketball-reference.com

## Centers
center_2001 <- rnorm(n = 100, mean = 0.83, sd = .3)
center_2021 <- rnorm(n = 100, mean = 1.6, sd = .4)

## Point Guards
pg_2001 <- abs(rnorm(n = 100, mean = 4.06, sd = 2)) ## abs so I can increase sd without getting negatives
pg_2021 <- abs(rnorm(n = 100, mean = 4.05, sd = 2))

## Shooting Guards
sg_2001 <- rnorm(n= 100, mean = 2.6, sd = .8)
sg_2021 <- rnorm(n = 100, mean = 2.2, sd = .7)

## Small Forwards
sf_2001 <- rnorm(n = 100, mean = 1.8, sd = .48)
sf_2021 <- rnorm(n = 100, mean = 1.9, sd = .5)

## Power Forwards
pf_2001 <- rnorm(n = 100, mean = 1.4, sd = .35)
pf_2021 <- rnorm(n = 100, mean = 1.8, sd = .45)
```  

## More Cleaning  
I needed additional transformations to graph the data as I intended.  

```{r dataframe, warning = FALSE, message = FALSE}

## Create dataframes from vectors


### Dataframes with no melt

## 2001 Assists
assists_2001 <- data.frame(center_2001, pg_2001, sg_2001, sf_2001, pf_2001)

## Rename Variables
assists_2001 <- assists_2001 %>% 
  rename(
    Center = center_2001,
    Point_Guard = pg_2001,
    Shooting_Guard = sg_2001,
    Shooting_Forward = sf_2001,
    Power_Forward = pf_2001
  )

## 2021 Assists
assists_2021 <- data.frame(center_2021, pg_2021, sg_2021, sf_2021, pf_2021)

## Rename Variables
assists_2021 <- assists_2021 %>% 
  rename(
    Center = center_2021,
    Point_Guard = pg_2021,
    Shooting_Guard = sg_2021,
    Shooting_Forward = sf_2021,
    Power_Forward = pf_2021
  )


## 2001 Assists Dataframe Melted
nba_assists_2001 <- melt(assists_2001)
colnames(nba_assists_2001) <- c('Position', 'APG')



## 2021 Assists Dataframe Melted
nba_assists_2021 <- melt(assists_2021)
colnames(nba_assists_2021) <- c('Position', 'APG')



```


## Graphical Analysis  

I ran a pairs.panel on unaltered dataset to look for any correlations, and view the general distribution.  There were no interesting associations of note.  From here, I wanted to visualize the differences with a boxplot sequence.  In this, there are three `Positions` that appear to have significant movement from 2001 to 2021; `Center`, `Shooting Guard`, and `Power Foward`.  `Point Guard` also seems to have increased over the years in variance.  
Additionally, I tallied the Assist Totals and multiplied them by 82 games (Season length), to get the Season totals for Assists.  This is represented in the bar chart, which demonstrates a rise in assists from 2001 to 2021.  


```{r analyze, message = FALSE, warning = FALSE}
## Assists 2001
pairs.panels(assists_2001)

## Assists 2021
pairs.panels(assists_2021)


# 2001 Boxplot
g1 <- ggplot(data = nba_assists_2001,
       mapping = aes(x = Position, y = APG, fill = Position))+
  geom_boxplot(alpha = .9, color = 'grey')+
  scale_fill_futurama()+
  labs(title = 'Assist Per Game by Position: 2001')+
  guides(fill = FALSE)+
  ylim(0,10)+
  theme_modern_rc()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 2021 Boxplot
g2 <- ggplot(data = nba_assists_2021,
       mapping = aes(x = Position, y = APG, fill = Position))+
  geom_boxplot(alpha = .9, color = 'grey')+
  scale_fill_futurama()+
  labs(title = 'Assist Per Game by Position: 2021')+
  guides(fill = FALSE)+
  ylim(0,10)+
  theme_modern_rc()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 2001 Boxplot Print
g1

# 2021 Boxplot Print
g2


## Total Assist Comparison
total_2001 <- sum(nba_assists_2001$APG * 82)
total_2021 <- sum(nba_assists_2021$APG * 82)

total_assists <- data.frame(total_2001, total_2021)

nba_total_assists <- melt(total_assists)
colnames(nba_total_assists) <- c('Year', 'Total')

ggplot(data = nba_total_assists,
       mapping = aes(x = Year, y = Total, fill = Year))+
  geom_col(position = 'dodge', color = 'black', size = 1, alpha = .9)+
  guides(fill = FALSE)+
  scale_fill_startrek()+
  theme_ipsum_tw()+
  labs(title = "Total Assists: 2001 vs 2021")+
  scale_x_discrete(labels = c('2001', '2021'))
```

## Further Cleaning  
More cleaning to set up t.tests
```{r stattest}

## 2001 vs 2021 Positions

## Center
center_data <- assists_2001 %>% 
  select(Center)
center_data[2] <- assists_2021 %>% 
  select(Center)

center_data <- center_data %>% 
  rename(
    C_2001 = Center,
    C_2021 = Center.1)

## PG
pg_data <- assists_2001 %>% 
  select(Point_Guard)
pg_data[2] <- assists_2021 %>% 
  select(Point_Guard)

pg_data <- pg_data %>% 
  rename(
    PG_2001 = Point_Guard,
    PG_2021 = Point_Guard.1)


## SG
sg_data <- assists_2001 %>% 
  select(Shooting_Guard)
sg_data[2] <- assists_2021 %>% 
  select(Shooting_Guard)

sg_data <- sg_data %>% 
  rename(
    SG_2001 = Shooting_Guard,
    SG_2021 = Shooting_Guard.1)


## SF
sf_data <- assists_2001 %>% 
  select(Shooting_Forward)
sf_data[2] <- assists_2021 %>% 
  select(Shooting_Forward)

sf_data <- sf_data %>% 
  rename(
    SF_2001 = Shooting_Forward,
    SF_2021 = Shooting_Forward.1)


## PF
pf_data <- assists_2001 %>% 
  select(Power_Forward)
pf_data[2] <- assists_2021 %>% 
  select(Power_Forward)

pf_data <- pf_data %>% 
  rename(
    PF_2001 = Power_Forward,
    PF_2021 = Power_Forward.1)

```

## t.test  
We saw there was a visual difference in some of the assists in the boxplots, but was there a statistically significant difference?  To test this, I ran a t.test; conditions are (mostly) normal (I have abs() on point_guard to prevent negative outliers).  The t.test confirmed the visual analysis; there was a statistically significant difference in the means in Assists per Game for `Center`, `Shooting_Guard`, and `Power_Forward`.  The direction of the difference aligns with our hypothesis - it would appear that Assists per Game have increased in the power positions.  


```{r asdf}

## Center t.test
t.test(center_data$C_2001, center_data$C_2021)

## PG t.test
t.test(pg_data$PG_2001, pg_data$PG_2021)

## SG t.test
t.test(sg_data$SG_2001, sg_data$SG_2021)

## SF t.test
t.test(sf_data$SF_2001, sf_data$SF_2021)

## PF t.test
t.test(pf_data$PF_2001, pf_data$PF_2021)
```   

## Change the Numbers  

```{r changer, warning = FALSE, message = FALSE}
## Change the numbers to alter results

## Centers
center_2001 <- rnorm(n = 10, mean = 0.83, sd = .3)
center_2021 <- rnorm(n = 10, mean = 2.6, sd = .4)

## Point Guards
pg_2001 <- abs(rnorm(n = 100, mean = 4.06, sd = 2)) ## abs so I can increase sd without getting negatives
pg_2021 <- abs(rnorm(n = 100, mean = 5.05, sd = 4))

## Shooting Guards
sg_2001 <- rnorm(n= 200, mean = 2.6, sd = .8)
sg_2021 <- rnorm(n = 200, mean = 2.2, sd = .7)

## Small Forwards
sf_2001 <- rnorm(n = 100, mean = 5.8, sd = .48)
sf_2021 <- rnorm(n = 100, mean = 1.9, sd = .5)

## Power Forwards
pf_2001 <- rnorm(n = 50, mean = 1.6, sd = .35)
pf_2021 <- rnorm(n = 50, mean = 1.8, sd = .45)


## Reload All Transformations


### Dataframes with no melt

## 2001 Assists
assists_2001 <- data.frame(center_2001, pg_2001, sg_2001, sf_2001, pf_2001)

## Rename Variables
assists_2001 <- assists_2001 %>% 
  rename(
    Center = center_2001,
    Point_Guard = pg_2001,
    Shooting_Guard = sg_2001,
    Shooting_Forward = sf_2001,
    Power_Forward = pf_2001
  )

## 2021 Assists
assists_2021 <- data.frame(center_2021, pg_2021, sg_2021, sf_2021, pf_2021)

## Rename Variables
assists_2021 <- assists_2021 %>% 
  rename(
    Center = center_2021,
    Point_Guard = pg_2021,
    Shooting_Guard = sg_2021,
    Shooting_Forward = sf_2021,
    Power_Forward = pf_2021
  )


## 2001 Assists Dataframe Melted
nba_assists_2001 <- melt(assists_2001)
colnames(nba_assists_2001) <- c('Position', 'APG')



## 2021 Assists Dataframe Melted
nba_assists_2021 <- melt(assists_2021)
colnames(nba_assists_2021) <- c('Position', 'APG')


## 2001 vs 2021 Positions

## Center
center_data <- assists_2001 %>% 
  select(Center)
center_data[2] <- assists_2021 %>% 
  select(Center)

center_data <- center_data %>% 
  rename(
    C_2001 = Center,
    C_2021 = Center.1)

## PG
pg_data <- assists_2001 %>% 
  select(Point_Guard)
pg_data[2] <- assists_2021 %>% 
  select(Point_Guard)

pg_data <- pg_data %>% 
  rename(
    PG_2001 = Point_Guard,
    PG_2021 = Point_Guard.1)


## SG
sg_data <- assists_2001 %>% 
  select(Shooting_Guard)
sg_data[2] <- assists_2021 %>% 
  select(Shooting_Guard)

sg_data <- sg_data %>% 
  rename(
    SG_2001 = Shooting_Guard,
    SG_2021 = Shooting_Guard.1)


## SF
sf_data <- assists_2001 %>% 
  select(Shooting_Forward)
sf_data[2] <- assists_2021 %>% 
  select(Shooting_Forward)

sf_data <- sf_data %>% 
  rename(
    SF_2001 = Shooting_Forward,
    SF_2021 = Shooting_Forward.1)


## PF
pf_data <- assists_2001 %>% 
  select(Power_Forward)
pf_data[2] <- assists_2021 %>% 
  select(Power_Forward)

pf_data <- pf_data %>% 
  rename(
    PF_2001 = Power_Forward,
    PF_2021 = Power_Forward.1)

```    

## Altered t.test   
Changing the numbers drastically effects the t.test, which was to be expected.  
```{r altertest}

## Center t.test
t.test(center_data$C_2001, center_data$C_2021)

## PG t.test
t.test(pg_data$PG_2001, pg_data$PG_2021)

## SG t.test
t.test(sg_data$SG_2001, sg_data$SG_2021)

## SF t.test
t.test(sf_data$SF_2001, sf_data$SF_2021)

## PF t.test
t.test(pf_data$PF_2001, pf_data$PF_2021)
```   
